\begin{authorcontributions}

I was responsible for software development, data processing, and analysis of all findings, and drafting manuscripts. 
Tristan Glatard was responsible for supervising and supporting all of my contributions.
The contributions of authors to each publication are described below.

\textbf{C.I – File-based localization of numerical perturbations in data analysis pipelines~\cite{salari2020spot}}

I was responsible for the tool development, data processing, analysis, drafting the manuscript, and designing the figures.
Lindsay B. Lewis and Alan C. Evans provided input on the dataset and pipelines, reviewed the results, and approved the final version of the manuscript.
Gregory Kiar and Tristan Glatard supported development processes and data visualization.
Tristan Glatard edited the manuscript, contributed to the interpretation of results, and supervised the findings of this work.

\textbf{C.II – Accurate simulation of operating system updates in neuroimaging using Monte-Carlo arithmetic~\cite{salari2021accurate}}

I was responsible for the software implementation, data processing, analysis, drafting the manuscript, and designing the figures.
All authors contributed to the editing of the manuscript, experimental design and discussed the results.
Yohan Chatelain helped with Monte-Carlo arithmetic simulations and software testing.
Gregory Kiar and Tristan Glatard provided software development support.
Tristan Glatard supervised the findings of this work.

\textbf{C.III – Software variability in fMRI analysis: comparing between-tool and numerical errors}

I was responsible for reproducing the experiments, data processing, drafting the manuscript, and designing the figures.
%X, Y supported the implementation of fMRI analyses and provided valuable feedback.
Gregory Kiar, Yohan Chatelain, and Tristan Glatard contributed to the experimental design and interpretation of results.
Tristan Glatard edited the manuscript and supervised the findings of this work.

\end{authorcontributions}